---
title: "Record logs in shiny apps"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{shinylogs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

```{r setup}
library(shinylogs)
```


## How to use

Add in your server : 

```{r}
# server
track_usage(storage_mode = store_json(path = "logs/"))
```

It will store logs in a sub-folder `logs/` of your application.

If needed (in complex application), you can force dependencies in UI with :

```{r}
# UI
use_tracking()
```



## Storage modes

Four modes are available:

* `storage_json()` : store logs has separate JSON files (one by session).
* `storage_rds()` : store logs has individually RDS files (one by session).
* `storage_sqlite()` : store logs in a SQLite database.
* `storage_json()` : don't store logs, print them in R console.

On a server, if you want to save logs on disk, don't forget to set right permission on the folder you want to save logs.


## Recorded informations

### Session

Those data are metadata about the application and the user's browser, here are the filed recorded : 

* **app** : name of the application
* **user** : name of the user (if using Shiny-server pro for example)
* **server_connected** : timestamp of when application has been launched (server time)
* **sessionid** : a session ID to match the session with other recorded informations (inputs, outputs, errors)
* **server_disconnected** : timestamp of when the application was disconnected (server time)
* **user_agent** : browser user-agent
* **screen_res** : resolution of the user screen (width x height)
* **browser_res** : resolution of the user browser (width x height)
* **pixel_ratio** : pixel ratio of the browser
* **browser_connected** : timestamp of when application has been launched (browser time, depends on user timezone)



### Inputs

Data about inputs, by default all inputs are recorded (even those not define by developper, like with {htmlwidgets} : {DT}, {leaflet}, ...)

* **sessionid** : the same ID as in session object
* **name** : inputId of the input
* **timestamp** : timestamp when the input has changed
* **value** : the value taken by the input (can be a list in case of complex input)
* **type** : type of input (if defined)
* **binding** : binding for the input (if defined)


### Outputs

Data recorded each time an output is refreshed :

* **sessionid** : the same ID as in session object
* **name** : outputId of the output
* **timestamp** : timestamp when the output has been updated
* **type** : type of output (if defined)
* **binding** : binding for the output (if defined)


### Errors

Errors are recorded only when propagated through an output, this is the red message users see in application, infos saved are:

* **sessionid** : the same ID as in session object
* **name** : outputId of the output where an error happened
* **timestamp** : timestamp of the error
* **error** : error message (if any)
* **value** : additional data for the error (generally NULL)




